\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual Multi-Loop Integrands

### 2023-09-21 notes/2loopNLSM.tex
push small change

The final constraints comes from ensuring that the ansatz correctly
reproduces the generalized unitarity cuts of the desired theory, thus
guaranteeing that the numerators encode the correct physical,
theory-dependent information.  Generalized unitarity methods has been
reviewed in many places so we will be light on details here
\cite{BCFUnitarity, Bern:2015ooa, Bourjaily:2017wjl,
  JJHenrikReview, FivePointN4BCJ, Bern:2012uf, BCJreview,
  Edison:2022jln, Edison:2022smn}, which was recently optimized for effective field theories like NLSM \cite{Carrasco:2023qgz}. Generalized unitarity equates products of on-shell tree amplitudes,
encoded via a graph $\gamma$, to sums over compatible\footnote{A
  diagram $g$ is compatible with the cut diagram $\gamma$ if $g$ is
  isomorphic to $\gamma$ or to any of the additional factorization
  channels of $\gamma$. } diagram numerators evaluated on the support
of the on-shell conditions and weighted by their uncut propagators
\begin{equation}
\label{eq:SchematicCutEquation}
\text{Cut}[\gamma]= \sum_{\substack{\text{states}\\\text{crossing } E(\gamma)}} \prod_{v \in V(\gamma)} \atree(v)
= \sum_{g \text{ compat } \gamma}
\frac{N[g] \vert_\text{cut}}{\prod \limits_{\text{uncut}} \ell^2}
 \, ,
\end{equation}
where $V(\gamma)$ are the vertices of $\gamma$ and $E(\gamma)$ its
edges.  The sum over states hides much of the complexity of cut
construction: for scalars the process is trivial while spinning states
greatly complicate matters (see
Refs.~\cite{Kosmopoulos:2020pcd,Edison:2022jln} for recent discussion
and detailed examples of evaluating state sums).  As an example of the
two types of expressions appearing in \cref{eq:SchematicCutEquation},
consider the so-called iterated two-particle cut which can be
represented as a product of trees via
\begin{equation}
  \label{eq:example-cut-prod}
  \ExCutLabeled{1}{2}{3}{4} = \!\!\!\!\! \sum_{\substack{\text{states of }\\ \{\ell_1, \ell_2, \ell_3, \ell_4\}}} \!\!\!\!\!\!\!
  \atree(1,2,\ell_1,\ell_2) \atree(-\ell_1,-\ell_2,\ell_3,\ell_4) \atree(-\ell_3,-\ell_4,3,4)
\end{equation}
or as a collection of diagram numerators via
\begin{align}
  \label{eq:example-cut-sum}
  \ExCutLabeled{1}{2}{3}{4} &=
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}{(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7})^2}
  \ N\left[
    \IUCContrib{1}{4}{3}{2}{\ell_5}{\ell_6}{\ell_7}
                              \right] \notag \\
  &\quad + 
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}
    {(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7'})^2}
  \ N\left[
    \IUCContrib{1}{3}{4}{2}{\ell_5}{\ell_6}{\ell_7'}
    \right] + 25\text{ others.}
\end{align}
When drawing cut diagrams, we will use gray blobs as in the left-hand
sides of \cref{eq:example-cut-prod,eq:example-cut-sum} and all drawn
legs are assumed to be on-shell.  When drawing numerator contributions
to cuts, we will not use blobs on vertices, and will use
\textcolor{cut-color}{dashed lines} bisecting edges to denote cut
lines and use colored edges to highlight the
\textcolor{ucp-color}{uncut propagators}.  From now on we will leave
the cut $\delta(\ell^2)$s implicit.

@@

\ace{Lets tighten up this paragraph a bit, especially the second half}
It is tempting to try to match the pion four-point two-loop integrand
to ZM on the cuts, but given that the cuts only probe the on-shell
four-point amplitude, which is either subtle or vanishes for ZM, it is
better to compare the off-shell numerators directly.  As local
functions, the numerators never suffer from any of the subtleties of
on-shell 2D kinematics.  Since the physical cuts are difficult to
compare, ZM theory is being used here simply as a tool to eliminate
generalized gauge freedom; ZM theory has the right power counting and
kinematic structure in 2D to compare against.  Since ZM theory is only
being used for the kinematical structure of its numerators, it is
natural to introduce a proportionality constant $z$ between the NLSM
and ZM numerators $n_\text{NLSM} \vert_{2D} = z ~ n_\text{ZM}$.  The
parameter $z$, which would normally be fixed on cuts, effectively
encodes the discrepancy between the coupling constants of the two
theories.\footnote{Floating the proportionality parameter $z$ is also
  motivated because ZM is believed to be quantum mechanically
  different from NLSM even though they are dual classically in two
  dimensions \cite{Nappi:1979ig}.  When the four-point
  pion amplitude is normalized to $-k_1\cdot k_2$ and the ZM vertex is
  normalized to $k_1^\mu \varepsilon_{\mu\nu}k_2^\nu$, the parameter
  $z$ is fixed to $216/565$, which is clearly unrelated to the
  difference in coupling constants.}  This introduces an extra degree of
freedom into the problem.  Mechanically, the pion numerator is matched
to ZM by first taking every possible assignment of right and left
movers for the external particles and then restricting the pion
numerator to 2D.  The loop momenta are restricted to 2D but left
off-shell.  After performing the 2D matching to the following ZM basis numerators:
\be
\ZMBasisDoubleBox =\left[\begin{split}
  \langle (l_1+k_4) l_1 \rangle\langle l_1l_2 \rangle\langle l_2 (l_2+k_1) \rangle  \langle (l_2+k_1)(l_2+k_1+k_2) \rangle 
\\
\langle (l_1+k_3+k_4)(l_2+k_1+k_2) \rangle\langle (l_1+k_3+k_4)(l_1+k_4) \rangle
\end{split}\right]
\ee
\be
\ZMBasisPentaTriangle =\left[\begin{split}
  \langle (l_1+k_4) l_1 \rangle\langle l_1l_2 \rangle\langle l_2 (l_2+k_1) \rangle  \langle (l_2+k_1)(l_1-k_1) \rangle 
\\
\langle (l_1+k_3+k_4)(l_1-k_1) \rangle\langle (l_1+k_3+k_4)(l_1+k_4) \rangle
\end{split}\right]
\ee
there are only 365 parameters of generalized gauge freedom. An overview of ZM theory and its color-dual construction can be found at the end of \cref{sec:CKLagrangians}. The loop momentum structure of ZM theory provides one final hint for
simplifying the pion numerators.  A generic term in the ZM numerator
(of either basis graph) looks schematically like
\begin{equation}
\label{eq:ZMLoopPowCount}
\ell_1^m \ell_2^n k^{12-(m+n)} \text{ where } 5 \leq m+n \leq 8,
\end{equation}
whereas a generic set of local, cubic Feynman rules could have
produced terms with up to $m+n=12$ powers of loop momenta.  When the
pion numerator is forced to have the loop power counting structure in
\cref{eq:ZMLoopPowCount} of ZM theory, the number of generalized gauge
freedom parameters reduces to 58.  The pion numerator appearing in the
ancillary files makes exactly this choice.

@@

\section{Conclusions and Outlook}\label{conclusions}
There is overwhelming evidence that YM and NLSM obey color-kinematics
duality at loop level, even though naive Lagrangians for these theories
do not manifest the duality.  At tree level, the duality can be
manifested by breaking manifest Bose symmetry at the Lagrangian level,
either by explicitly picking different gluon states (as in SDYM) or by
picking states that only restore Bose symmetry in the final amplitude
(as in NLSM).  All of these theories can be encapsulated in the
semi-abelian YM theory presented in \cref{eq:semiYM} which obscures
the Bose symmetry of the external states due to the presence of the
two gauge fields $A_\mu$ and $\bar{A}_\mu$.  The kinetic mixing
between the two gauge fields permits globally color-dual 1-loop
integrands, but special care must be taken so that the integrands
produced from semi-abelian YM correspond to SDYM or NLSM.  Due to the
kinetic mixing that stems from the broken Bose symmetry, it is
impossible to construct a two-loop integrand for semi-abelian YM theory,
that is, the theory is one-loop exact.  Obtaining a globally color-dual
integrand beyond one-loop then requires sacrificing more than just Bose
symmetry and in this paper it is argued that off-shell locality must
be abandoned.\footnote{Locality is relaxed in \cite{Mogull:2015adi} in
  order to obtain color-dual two-loop integrands.}  \ace{I want to
  rephrase especially the last sentence.}

\jm{I am intentionally shying away from really firm statements like
  `you must abandon locality' because there are always loopholes and
  we made a lot of other assumptions in fixing our \ansatze{}.}

Kinematic ``numerators'' that are rational functions of the kinematics
appear unnatural from the perspective of point-like quantum field
theories. After all, operators that produce rational functions of
kinematics must be non-local.\ace{A few more connecting sentences
  here.  This is a bit of a big jump without some amount of summary}
Thus, in order to identify valid color-dual representation of
Yang-Mills beyond tree-level, we must relax the constraint of
\textit{off-shell locality}, and realize locality only in the limit of
\textit{on-shell} kinematics. Below we provide an overview of what one
might expect for such a construction, and provide some simple examples
at tree-level that manifest this structure.

>>>>

The final constraints comes from ensuring that the ansatz correctly
reproduces the generalized unitarity cuts of the desired theory, thus
guaranteeing that the numerators encode the correct physical,
theory-dependent information.  Generalized unitarity methods has been
reviewed in many places so we will be light on details here
\cite{BCFUnitarity, Bern:2015ooa, Bourjaily:2017wjl,
  JJHenrikReview, FivePointN4BCJ, Bern:2012uf, BCJreview,
  Edison:2022jln, Edison:2022smn}, which was recently optimized for effective field theories like NLSM \cite{Carrasco:2023qgz}. Generalized unitarity equates products of on-shell tree amplitudes,
encoded via a graph $\gamma$, to sums over compatible\footnote{A
  diagram $g$ is compatible with the cut diagram $\gamma$ if $g$ is
  isomorphic to $\gamma$ or to any of the additional factorization
  channels of $\gamma$. } diagram numerators evaluated on the support
of the on-shell conditions and weighted by their uncut propagators
\begin{equation}
\label{eq:SchematicCutEquation}
\text{Cut}[\gamma]= \sum_{\substack{\text{states}\\\text{crossing } E(\gamma)}} \prod_{v \in V(\gamma)} \atree(v)
= \sum_{g \text{ compat } \gamma}
\frac{N[g] \vert_\text{cut}}{\prod \limits_{\text{uncut}} \ell^2}
 \, ,
\end{equation}
where $V(\gamma)$ are the vertices of $\gamma$ and $E(\gamma)$ its
edges.  The sum over states hides much of the complexity of cut
construction: for scalars the process is trivial while spinning states
greatly complicate matters (see
Refs.~\cite{Kosmopoulos:2020pcd,Edison:2022jln} for recent discussion
and detailed examples of evaluating state sums).  As an example of the
two types of expressions appearing in \cref{eq:SchematicCutEquation},
consider the so-called iterated two-particle cut which can be
represented as a product of trees via
\begin{equation}
  \label{eq:example-cut-prod}
  \ExCutLabeled{1}{2}{3}{4} = \!\!\!\!\! \sum_{\substack{\text{states of }\\ \{\ell_1, \ell_2, \ell_3, \ell_4\}}} \!\!\!\!\!\!\!
  \atree(1,2,\ell_1,\ell_2) \atree(-\ell_1,-\ell_2,\ell_3,\ell_4) \atree(-\ell_3,-\ell_4,3,4)
\end{equation}
or as a collection of diagram numerators via
\begin{align}
  \label{eq:example-cut-sum}
  \ExCutLabeled{1}{2}{3}{4} &=
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}{(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7})^2}
  \ N\left[
    \IUCContrib{1}{4}{3}{2}{\ell_5}{\ell_6}{\ell_7}
                              \right] \notag \\
  &\quad + 
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}
    {(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7'})^2}
  \ N\left[
    \IUCContrib{1}{3}{4}{2}{\ell_5}{\ell_6}{\ell_7'}
    \right] + 25\text{ others.}
\end{align}
When drawing cut diagrams, we will use gray blobs as in the left-hand
sides of \cref{eq:example-cut-prod,eq:example-cut-sum} and all drawn
legs are assumed to be on-shell.  When drawing numerator contributions
to cuts, we will not use blobs on vertices, and will use
\textcolor{cut-color}{dashed lines} bisecting edges to denote cut
lines and use colored edges to highlight the
\textcolor{ucp-color}{uncut propagators}.  From now on we will leave
the cut $\delta(\ell^2)$ factors implicit.

@@

%\ace{Lets tighten up this paragraph a bit, especially the second half}

It is tempting to try to match the pion four-point two-loop integrand
to ZM on the cuts, but given that the cuts only probe the on-shell
four-point amplitude, which is either subtle or vanishes for ZM, it is
better to compare the off-shell numerators directly.  As local
functions, the numerators never suffer from any of the subtleties of
on-shell 2D kinematics.  Since the physical cuts are difficult to
compare, ZM theory is being used here simply as a tool to eliminate
generalized gauge freedom; ZM theory has the right power counting and
kinematic structure in 2D to compare against.  Since ZM theory is only
being used for the kinematical structure of its numerators, it is
natural to introduce a proportionality constant $z$ between the NLSM
and ZM numerators $n_\text{NLSM} \vert_{2D} = z ~ n_\text{ZM}$.  The
parameter $z$, which would normally be fixed on cuts, effectively
encodes the discrepancy between the coupling constants of the two
theories.\footnote{Floating the proportionality parameter $z$ is also
  motivated because ZM is believed to be quantum mechanically
  different from NLSM even though they are dual classically in two
  dimensions \cite{Nappi:1979ig}.  When the four-point
  pion amplitude is normalized to $-k_1\cdot k_2$ and the ZM vertex is
  normalized to $k_1^\mu \varepsilon_{\mu\nu}k_2^\nu$, the parameter
  $z$ is fixed to $216/565$, which is clearly unrelated to the
  difference in coupling constants.}  This introduces an extra degree of
freedom into the problem.  Mechanically, the pion numerator is matched
to ZM by first taking every possible assignment of right and left
movers for the external particles and then restricting the pion
numerator to 2D.  The loop momenta are restricted to 2D but left
off-shell.  After performing the 2D matching to the following ZM basis numerators:
\be
\ZMBasisDoubleBox =\left[\begin{split}
  \langle (l_1+k_4) l_1 \rangle\langle l_1l_2 \rangle\langle l_2 (l_2+k_1) \rangle  \langle (l_2+k_1)(l_2+k_1+k_2) \rangle 
\\
\langle (l_1+k_3+k_4)(l_2+k_1+k_2) \rangle\langle (l_1+k_3+k_4)(l_1+k_4) \rangle
\end{split}\right]
\ee
\be
\ZMBasisPentaTriangle =\left[\begin{split}
  \langle (l_1+k_4) l_1 \rangle\langle l_1l_2 \rangle\langle l_2 (l_2+k_1) \rangle  \langle (l_2+k_1)(l_1-k_1) \rangle 
\\
\langle (l_1+k_3+k_4)(l_1-k_1) \rangle\langle (l_1+k_3+k_4)(l_1+k_4) \rangle
\end{split}\right]
\ee
there are only 365 parameters of generalized gauge freedom. An overview of ZM theory and its color-dual construction can be found at the end of \cref{sec:CKLagrangians}. The loop momentum structure of ZM theory provides one final hint for
simplifying the pion numerators.  A generic term in the ZM numerator
(of either basis graph) looks schematically like
\begin{equation}
\label{eq:ZMLoopPowCount}
\ell_1^m \ell_2^n k^{12-(m+n)} \text{ where } 5 \leq m+n \leq 8,
\end{equation}
whereas a generic set of local, cubic Feynman rules could have
produced terms with up to $m+n=12$ powers of loop momenta.  When the
pion numerator is forced to have the loop power counting structure in
\cref{eq:ZMLoopPowCount} of ZM theory, the number of generalized gauge
freedom parameters reduces to 58.  The pion numerator appearing in the
ancillary files makes exactly this choice.

@@

\section{Conclusions and Outlook}\label{conclusions}
There is overwhelming evidence that YM and NLSM obey color-kinematics
duality at loop level, even though naive Lagrangians for these theories
do not manifest the duality.  At tree level, the duality can be
manifested by breaking manifest Bose symmetry at the Lagrangian level,
either by explicitly picking different gluon states (as in SDYM) or by
picking states that only restore Bose symmetry in the final amplitude
(as in NLSM).  All of these theories can be encapsulated in the
semi-abelian YM theory presented in \cref{eq:semiYM} which obscures
the Bose symmetry of the external states due to the presence of the
two gauge fields $A_\mu$ and $\bar{A}_\mu$.  The kinetic mixing
between the two gauge fields permits globally color-dual 1-loop
integrands, but special care must be taken so that the integrands
produced from semi-abelian YM correspond to SDYM or NLSM.  Due to the
kinetic mixing that stems from the broken Bose symmetry, it is
impossible to construct a two-loop integrand for semi-abelian YM theory,
that is, the theory is one-loop exact.  Obtaining a globally color-dual
integrand beyond one-loop then requires sacrificing more than just Bose
symmetry and in this paper it is argued that off-shell locality must
be abandoned.\footnote{Locality is relaxed in \cite{Mogull:2015adi} in
  order to obtain color-dual two-loop integrands.}

Kinematic ``numerators'' that are rational functions of the kinematics
appear unnatural from the perspective of point-like quantum field
theories. After all, operators that produce rational functions of
kinematics must be non-local. Thus, in order to identify valid color-dual representation of
Yang-Mills beyond tree-level, we must relax the constraint of
\textit{off-shell locality}, and realize locality only in the limit of
\textit{on-shell} kinematics. Below we provide an overview of what one
might expect for such a construction, and provide some simple examples
at tree-level that manifest this structure.
